---
title: "Organic Geochemistry Analysis"
author: "Adam McCarthy"
date: "29 august 2017"
output: html_document
---

Dataset © Commonwealth of Australia (Geoscience Australia) 2017

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(ggthemes)
library(scales)
library(gridExtra)
library(GGally)
```

Taking data provided by Geoscience Australlia allows for an exploratory
data analysis to be undertaken on geochemical rock eval. data.

This project is a template to allow for anyone to try out R using a
large dataset of geochemical data points.

```{r import_data, echo=FALSE, warning=FALSE}
# Data is stored one folder up, then within the data folder.
setwd("..")
f_path <- file.path("data", "organic_geochem_tab.csv")

# Read in the tab deliminated version to format properly for R.
df_gc <- read.csv(f_path, sep = "\t",  # Tab seperated file
               encoding = "UTF-8",  # File encoding required to read
               stringsAsFactors = FALSE,  # Manually assign factors after import
               na.strings = c("","NA"))  # Specify that blank strings are NAN

# Remove X as this is the same as index
df_gc$X <- NULL
```

Example of plots.

Univariate plots:

```{r echo=FALSE, warning=FALSE, U2, fig.height = 3, fig.width = 5, fig.align="center"}
small_hist <- function(data_v, name, scale, df=df_gc) {
plt <- ggplot(data = df, aes(data_v)) + 
  geom_histogram(bins = 50,
                 col = "white",
                 fill = "#9099A2"
                 ) +
  theme_tufte() +
  labs(
    title = name
    ) +
  labs(
    x = scale,
    y = "Count"
    )

return(plt)
}

plt1 <- small_hist(df_gc$TOC.wt..., 
  "TOC", "%")

# Convert to a logarithmic scale to acount for the distribution of data points.
plt1 <- plt1 + scale_x_log10()

plt1
```

Bivariate plots:

```{r echo=FALSE, warning=FALSE, scatter_sulf, fig.height = 5, fig.width = 5}
ggplot(aes(y = df_gc$HI.mg.HC.g.TOC., x = df_gc$OI.mg.CO2.g.TOC.),
       data = df_gc) +
  geom_point(na.rm = FALSE,
             alpha = 0.2,
             position = position_jitter(),
             color = '#192231') +
  theme_tufte() +
  labs(
    title = "OI vs HI",
    y = "HI",
    x = "OI"
    ) + 
  xlim(0, 
       500
       ) +
  ylim(0, 
       1000
       )
```

Correlation matrix:

```{r echo=FALSE, correlation_matrix, fig.height = 5, fig.width = 5, message=FALSE, warning=FALSE}
ggcorr(df_gc)  +
  theme_tufte()
```

Pair plot:

```{r echo=FALSE, freq_pair_plot, fig.height = 9, fig.width = 9, message=FALSE, warning=FALSE,}
# List of columns to drop as they have too many string values for a pair plot
drops <- c("WELL.NAME", "SAMPLE.ID", "ORIGINATOR", 
           "COMMENTS", "ROCKEVAL_ANALYST", "VR_ANALYST")

ggpairs(df_gc[ , !(names(df_gc) %in% drops)]
        ) +
  theme_minimal()
```
